---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dccvalidator

[![Travis-CI Build Status](https://travis-ci.org/Sage-Bionetworks/dccvalidator.svg?branch=master)](https://travis-ci.org/Sage-Bionetworks/dccvalidator)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

Validate CNS Data and Metadata

## Installation

```{r installation, eval = FALSE}
devtools::install_github("Sage-Bionetworks/dccvalidator")
```

## Check annotations

You can check whether a Synapse file has valid annotation keys and values. You
can also check file views and data frames to see if their columns correspond to
valid annotations, and if the values in the columns are valid.

```{r check-annotations, eval = FALSE}
library("synapser")
library("dccvalidator")
synLogin()

## File
my_file <- synGet("syn17038065", downloadFile = FALSE)
check_annotation_keys(my_file)
check_annotation_values(my_file)

## File view
fv <- synTableQuery("SELECT * FROM syn17038067")
check_annotation_keys(fv)
check_annotation_values(fv)

## Data frame
dat <- data.frame(assay = "foo", b = 2)
check_annotation_keys(dat)
check_annotation_values(dat)
```

Please note that the dccvalidator project is released with a [Contributor Code
of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you
agree to abide by its terms.
