---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dccvalidator

[![Travis-CI Build Status](https://travis-ci.org/Sage-Bionetworks/dccvalidator.svg?branch=master)](https://travis-ci.org/Sage-Bionetworks/dccvalidator)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

dccvalidator is a package and Shiny app to perform data validation and QA/QC.
It's used in the [AMP-AD](https://ampadportal.org/) and
[PsychENCODE](http://www.psychencode.org) consortia to validate data prior to
data releases.

## Installation

```{r installation, eval = FALSE}
devtools::install_github("Sage-Bionetworks/dccvalidator")
```

## Check data

dccvalidator provides functions for checking the following common data quality
issues:

- Annotation keys and values conform to a controlled vocabulary
- Column names match those of an associated metadata template
- Certain columns are not empty
- Certain columns are complete
- Identifiers match between two metadata files (e.g. all individuals in one file
  are also present in another)
- Check that identifiers are unique within a file

# Data submission validation

This package contains a Shiny app to validate manifests and metadata for AMP-AD
studies. It uses the dccvalidator package to check for common data quality
issues and gives realtime feedback to the data contributor on errors that need
to be fixed. The reporting UI is heavily inspired by the 
[MetaDIG project's metadata quality reports](https://knb.ecoinformatics.org/quality/s=knb.suite.1/doi%3A10.5063%2FF12V2D1V).

The application also allows users to submit documentation of their study, a
description of the methods used, etc.

# Local development

`dccvalidator` uses pre-commit hooks to check for common issues, such as code
style (which should conform to [tidyverse style](https://style.tidyverse.org/)),
code parsability, and up-to-date .Rd documentation. To use, you will need to
install [pre-commit](https://pre-commit.com/#intro). If on a Mac, I recommend
using [homebrew](https://brew.sh/):

```
brew install pre-commit
```

Then, within this git repo, run:

```
pre-commit install
```

When you commit your changes, pre-commit will run the checks described above,
and the commit will fail if the checks do not pass. If you are experiencing
issues with the checks and want to commit your work and worry about them later,
you can run `git commit --no-verify` to skip all checks. Or, you can skip
certain hooks by their ID (as shown in the file `.pre-commit-config.yaml`), e.g.
`SKIP=roxygenize git commit -m "foo"`.

---

Please note that the dccvalidator project is released with a [Contributor Code
of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you
agree to abide by its terms.
