---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dccvalidator

[![Travis-CI Build Status](https://travis-ci.org/Sage-Bionetworks/dccvalidator.svg?branch=master)](https://travis-ci.org/Sage-Bionetworks/dccvalidator)
[![Coverage status](https://codecov.io/gh/Sage-Bionetworks/dccvalidator/branch/master/graph/badge.svg)](https://codecov.io/github/Sage-Bionetworks/dccvalidator?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

Validate CNS data/metadata and annotations.

## Installation

```{r installation, eval = FALSE}
devtools::install_github("Sage-Bionetworks/dccvalidator")
```

## Check annotations

You can check whether a Synapse file has valid annotation keys and values. You
can also check file views and data frames to see if their columns correspond to
valid annotations, and if the values in the columns are valid.

```{r login, message = FALSE, results = "hide"}
library("synapser")
library("dccvalidator")
synLogin()
```

```{r check-annotations }
## Get annotations
annots <- get_synapse_annotations()

## File
my_file <- synGet("syn17038065", downloadFile = FALSE)
check_annotation_keys(my_file, annots)
check_annotation_values(my_file, annots)

## File view
fv <- synTableQuery("SELECT * FROM syn17038067")
check_annotation_keys(fv, annots)
check_annotation_values(fv, annots)

## Data frame
dat <- data.frame(assay = "foo", b = 2)
check_annotation_keys(dat, annots)
check_annotation_values(dat, annots)
```

If you do not provide the annotations data frame to check against, the functions
will download the data automatically with `get_synapse_annotations()`.

```{r, check-annotations-2}
my_file <- synGet("syn17038065", downloadFile = FALSE)
check_annotation_keys(my_file)
check_annotation_values(my_file)
```

The result of the `check_*()` functions is a custom condition object. When the
check returns an invalid result, the problematic data is included.

```{r check-object}
dat <- data.frame(assay = "foo", b = 2)
res <- check_annotation_keys(dat)
res$data
```

If you instead wish to view which annotations _are_ valid, you can use
`valid_annotation_keys()` and `valid_annotation_values()`.

## Check template columns

You can check that the columns in your (meta)data conform to metadata templates.

```{r check-cols-assay}
dat <- data.frame(specimenId = "a", assay = "rnaSeq")
check_cols_assay(dat, template = "rnaSeq")
```

## Check specimen and individual IDs

To make sure the specimen IDs or individual IDs match between files (e.g. that
the specimens described in a biospecimen file match those in an assay metadata
file), `check_specimen_ids_match()` and `check_indiv_ids_match()`.

```{r check-ids}
## Two data frames with different specimen IDs
a <- data.frame(specimenID = LETTERS[1:3])
b <- data.frame(specimenID = LETTERS[1:4])
(result <- check_specimen_ids_match(a, b, "biospecimen", "assay"))
result$data
```

---

Please note that the dccvalidator project is released with a [Contributor Code
of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you
agree to abide by its terms.
