---
output:
  md_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# dccvalidator

[![Travis-CI Build Status](https://travis-ci.org/Sage-Bionetworks/dccvalidator.svg?branch=master)](https://travis-ci.org/Sage-Bionetworks/dccvalidator)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

Validate CNS data/metadata and annotations.

## Installation

```{r installation, eval = FALSE}
devtools::install_github("Sage-Bionetworks/dccvalidator")
```

## Check annotations

You can check whether a Synapse file has valid annotation keys and values. You
can also check file views and data frames to see if their columns correspond to
valid annotations, and if the values in the columns are valid.

```{r load, echo = FALSE}
pkgload::load_all()
```

```{r login, message = FALSE, results = "hide"}
library("synapser")
library("dccvalidator")
synLogin()
```

```{r check-annotations }
## Get annotations
annots <- get_synapse_annotations()

## File
my_file <- synGet("syn17038065", downloadFile = FALSE)
check_annotation_keys(my_file, annots)
check_annotation_values(my_file, annots)

## File view
fv <- synTableQuery("SELECT * FROM syn17038067")
check_annotation_keys(fv, annots)
check_annotation_values(fv, annots)

## Data frame
dat <- data.frame(assay = "foo", b = 2)
check_annotation_keys(dat, annots)
check_annotation_values(dat, annots)
```

If you do not provide the annotations data frame to check against, the functions
will download the data automatically with `get_synapse_annotations()`.

```{r, check-annotations-2}
my_file <- synGet("syn17038065", downloadFile = FALSE)
check_annotation_keys(my_file)
check_annotation_values(my_file)
```

The result of the `check_*()` functions is a custom condition object. When the
check returns an invalid result, the problematic data is included.

```{r check-object}
dat <- data.frame(assay = "foo", b = 2)
res <- check_annotation_keys(dat)
res$data
```

If you instead wish to view which annotations _are_ valid, you can use
`valid_annotation_keys()` and `valid_annotation_values()`.

## Check template columns

You can check that the columns in your (meta)data conform to metadata templates.

```{r check-cols-assay}
dat <- data.frame(specimenId = "a", assay = "rnaSeq")
check_cols_assay(dat, template = "rnaSeq")
```

## Check specimen and individual IDs

To make sure the specimen IDs or individual IDs match between files (e.g. that
the specimens described in a biospecimen file match those in an assay metadata
file), `check_specimen_ids_match()` and `check_indiv_ids_match()`.

```{r check-ids}
## Two data frames with different specimen IDs
a <- data.frame(specimenID = LETTERS[1:3])
b <- data.frame(specimenID = LETTERS[1:4])
(result <- check_specimen_ids_match(a, b, "biospecimen", "assay"))
result$data
```

# Data submission validation

This package contains a Shiny app to validate manifests and metadata for AMP-AD
studies. It uses the dccvalidator package to check for common data quality
issues and gives realtime feedback to the data contributor on errors that need
to be fixed. The reporting UI is heavily inspired by the 
[MetaDIG project's metadata quality reports](https://knb.ecoinformatics.org/quality/s=knb.suite.1/doi%3A10.5063%2FF12V2D1V).

## Deployment

This section is applicable to Sage employees who are deploying the application
on our Shiny Pro server. To learn about the server and how to get credentials,
please read the 
[Confluence documentation](https://sagebionetworks.jira.com/wiki/spaces/SageShinyServer/pages/75497489/Shiny+Server).

The app is deployed on the server through the following steps:

1. ssh into the Shiny Pro server and navigate to `/home/kwoo/ShinyApps/dccvalidator-app`
1. `git pull` changes from GitHub
1. To ensure packages are up-to-date, run `Rscript -e "renv::restore()"`

You may need to run `touch restart.txt` afterward to ensure the application is
restarted.

If you want to deploy the app in a different location (e.g. because you want to
stand up a new version that is customized for a different community):

1. ssh into the Shiny Pro server
1. Create a folder under `/home/yourusername/ShinyApps`
1. `git clone` the repository into the folder you've created
1. If needed, make any changes to the application's behavior by editing the
   files, or check out a branch that contains your changes
1. To ensure packages are up-to-date, run `Rscript -e "renv::restore()"`

Again, you may need to run `touch restart.txt` afterward to ensure the
application is restarted.

## Local development

`dccvalidator` uses pre-commit hooks to check for common issues, such as code
style (which should conform to [tidyverse style](https://style.tidyverse.org/)),
code parsability, and up-to-date .Rd documentation. To use, you will need to
install [pre-commit](https://pre-commit.com/#intro). If on a Mac, I recommend
using [homebrew](https://brew.sh/):

```
brew install pre-commit
```

Then, within this git repo, run:

```
pre-commit install
```

---

Please note that the dccvalidator project is released with a [Contributor Code
of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you
agree to abide by its terms.
